A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2019-01-10 article/01_Introduction.Rmd
and even more love for the introduction

Already the assumption that the cremation aspect of Early Bronze Age burial rituals in Britain can be directly compared to the same aspect in Hungary requires a theoretical foundation. *Cultural transmission theory* provides a framework where this claim can be explored as a hypothesis for further research. The spread of ideas in human populations over space and time is a major topic of *Cultural evolution theory* [@creanzaCulturalEvolutionaryTheory2017]. This line of thought is at least as old as theories of biological evolution and underwent a long and versatile research history from *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, against *Cultural relativism* and back to *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004]. In the late seventies, the discussion took another turn away from the political and social structure to the more general observation of human behaviour. Three major theoretical movements emerged in this context: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000]. The latter -- heavily popularised by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007]. Just like in biological evolution there might be selection and mutation within *Social learning* and therefore methods of *Population genetics* could be employed within cultural anthropology -- of course with the much faster pace of cultural exchange in mind. A lot of theoretical and quantitative research arose from these assumptions: e.g. @cavalli-sforzaCulturalTransmissionEvolution1981, @boydCultureEvolutionaryProcess1985, @teltserEvolutionaryArchaeologyMethodological1995 or @obrienEvolutionaryArchaeologyTheory1996.

>>>>

To answer these questions, a theoretical framework is required that allows to compare human behaviour across spatial, temporal and cultural boundaries: The assumption that the cremation aspect of Early Bronze Age burial rituals in Britain can be directly compared to the same aspect in Hungary is not self-evident. *Cultural evolution theory* provides such a framework [@creanzaCulturalEvolutionaryTheory2017]. The spread of ideas in human populations over space and time is one of it's major topics: *Cultural transmission*.

*Cultural evolution theory* is at least as old as theories of biological evolution and underwent a long and versatile research history. Its beginnings are in *Evolutionism* and *Social Darwinism* in the 19th and early 20th century, matured in conflict with *Cultural relativism* into *Neoevolutionism* up until the nineteen sixties and seventies [@petermannGeschichteEthnologie2004]. In the late seventies three major theoretical movements emerged: *Evolutionary psychology*, *Human behavioural ecology* and *Dual inheritance theory* [@smithThreeStylesEvolutionary2000]. The latter -- heavily popularised by Richard Dawkin's *Memetics* [@dawkinsSelfishGene1976] -- establishes the notion of culture acting as secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007]: Just like in biological evolution also *Social learning* might be prone to selection and mutation. Under this assumption methods of *Population genetics* could be employed within cultural anthropology to analyse the spread and development of ideas -- of course with the much faster pace of cultural exchange in mind. A lot of theoretical and quantitative research follows this line of thought: e.g. @cavalli-sforzaCulturalTransmissionEvolution1981, @boydCultureEvolutionaryProcess1985, @teltserEvolutionaryArchaeologyMethodological1995 or @obrienEvolutionaryArchaeologyTheory1996.
